\LARGE \bf Cascading Situation Assessment for Robots

### 2017-02-28 paper.tex
Update on Overleaf.

Software components accessing \uwds worlds are called clients. Some standard
clients (like a 3D visualization tool) are provided with the \uwds package (see
section~\ref{std_clients}). Clients are otherwise written by the end users
using the \uwds client API (see section~\ref{api}).

@@

Four specific types of clients can be distinguished: \textbf{root clients} which
create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
\emph{ROS interface} clients on Figure~\ref{fig|scene}); \textbf{leaf clients}
which on the contrary only read worlds, without modifying them (like the
\emph{Computation of spatial relations} client on Figure~\ref{fig|scene});
\textbf{filters} that copy an input world into an output world, performing
some filtering operation in-between (like the client \textit{Physics-based
position correction}); and \textbf{transformers} which transform one
representation into another (like the \emph{Semantic mapping} client on
Figure~\ref{fig|scene}).

@@

Each world is technically independent from every others. Dependencies between
worlds arise from the clients' connections. For instance, filters effectively
create a dependency between worlds. On Figure~\ref{fig|scene}, the \textit{Physics-based
position correction} client creates a dependency between the world {\tt base} (which
represents here the result of raw sensor fusion) and the world {\tt corrected}
which would be a physically-consistent copy of {\tt base}.
As a result, an \uwds network can be seen as a dependency graph between worlds (where
cyclic dependencies are permissible).

@@

Nodes in an \uwds scene graph have four possible types: \textbf{objects} that
represent concrete physical objects (typically with one or several associated 3D
meshes); \textbf{entities} that represent abstract entities like reference
frames or groups of objects; \textbf{perspectives} that represent viewpoints
onto the scene (like cameras or human gaze); and \textbf{fields} that represent
scalar or vector fields (like the visibility of an object, the working space of
robot, etc. -- note that fields are not yet implemented in the current version
of \uwds, see section~\ref{futurework}).

@@

Complementing the spatial representation encapsulated in the scene graph, each
world also store the world's \emph{timeline}. This data structure is shared
and synchronised amongst the clients in the same way as the scene graph.
Clients can record in the timeline both \emph{events} (duration-less states) and
\emph{situations}, \ie states with a start time and a (possibly open-ended) end time.

Importantly, the \uwds server automatically generates a snapshot of
the scene graph whenever an event or situation is added to the timeline. The
snapshot is associated to the event, which allows clients to effectively retrieve past
states of the world. This capability is thought to be primarly used by \uwds clients
performing action recognition.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{uwds-screenshot}
    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
    tool. In this particular example, the 3D meshes have been pre-loaded using
    {\tt uwds load}. Their positions are then updated at run time using the
    robot's sensors and proprioception (joint state).}
    \label{fig|uwds-view}
\end{figure}

@@

Critically, the network is essentially hidden to the client: from the user
perspective, the environment model is manipulated as a local data structure (see
Listing~\ref{lst|pythonapi}). Modifications to the model are asynchronously synchronised with
a central server (the {\tt underworlded} daemon) and broadcasted to every other
client connected to the same world.

As previously mentioned, worlds are composite data structures comprising of a
scene graph and a timeline. These data structures are synchronised using
Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
high throughput, reliability and cross-platform/cross-language support. The \uwds
API is specifically discussed hereafter, in section~\ref{api}.

\uwds is meant to broadcast complex environment representations (typically
including large geometric datasets, like meshes) in real-time. \uwds itself does
not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
fusion, physics simulation, etc. -- are performed by the clients themselves) and
as such, the performance bottleneck is essentially the network's data
throughput.  In that regard, one of the simple yet critical optimisation
performed by \uwds is automatic caching of mesh data. Mesh data are not
transmitted when nodes are updated; only a hash value of the mesh data. The
client can then request the full data whenever they actually need it.

@@

As mentioned, \uwds uses Google's gRPC as message passing protocol. The protocol
is explicitly defined (using the \emph{protocol
buffers}\footnote{\url{https://developers.google.com/protocol-buffers/}}
interface definition language), and bindings to various languages and platforms
can be automatically generated from the protocol definition file (as of Feb
2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
and Java on Windows, Mac and Linux).  The cross-platform/cross-language support
of gRPC is especially welcomed in the academic context, as it offers ease and
flexibility to plug a variety of pre-existing components into an \uwds network.

@@

\uwds currently offers such a high-level client library for Python only (a C++
library is under development).  Listing~\ref{lst|pythonapi} gives a complete
example of an \uwds client performing simple filtering: the client continuously
listen for changes in an input world, removes some objects (in this case, items
whose name contain the string ``{\tt wall}''), and forward all other changes to
an output world, effectively making the output world a copy of the input world
with all walls removed.

@@

        for node in in_word.scene.nodes:
            if("wall" not in node.name):
                out_world.scene.nodes.update(node)

\end{pythoncode}
    \caption{Example of a simple yet complete \uwds filter, written in Python:
    the client connects to the \uwds network, walk through the scene graph of
    world {\tt example\_in}, filter out some objects, and publishes the
    remaining objects in the world {\tt example\_out}.}

@@

Interestingly, while \uwds deals with 3D geometries and scenes, it does
represent 3D entities purely as data structures: no visual representation is
involved (and as such, the \uwds server and core libraries do not depend on any
graphics library like OpenGL). However, for all practical purposes, the ability
to visualise the content of a scene is desirable. \uwds provides a standard
client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
using OpenGL (Figure~\ref{fig|uwds-view}).

This tool also support basic object manipulations, that are broadcasted to the
other \uwds clients connected to the same world.

@@

Geometric data like meshes can be loaded into a world (or rather, into the whole
network as geometric data are treated as immutable resources shared amongst all
worlds) by simply sending an array of vertices and faces to the \uwds server.
As such, any client can load 3D data that is automatically made available to the
whole network. For example, an object recognition module recognises a new
object. It sends the mesh to the \uwds server and creates a new node which
references the mesh. At the next frame update, the 3D viewer sees a new node
referencing a yet unknown mesh. The viewer request the mesh data from the
server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
the node transformation is updated, the 3D viewer can re-render the node without
having to download the mesh from the server again.

Often, objects manipulated by the robot have known meshes with corresponding CAD
models that can be conveniently pre-loaded. In these case, \uwds provides a
tool, {\tt uwds load}, that loads a mesh into a \uwds network (and optionally,
creates a node) from a large range of 3D format (including Collada, FBX, OBJ,
Blender)\footnote{The underlying import
capability is provided by the {\sc assimp} library.
\url{http://assimp.sourceforge.net/}}.

@@

\uwds provides a range of tools to inspect a running the network. Graphical
tools ({\tt uwds explorer} and {\tt uwds timeline}, see Figure~\ref{fig|explorer})
provide an user-friendly overview of the system's graph with the connections
between the clients and the worlds, as well as their activity.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{tools}
    \caption{Screenshots of the timeline (top) and network (bottom)
    introspection tools. Worlds are represented as boxes, clients as ellipses.}
    \label{fig|explorer}
\end{figure}

Specialised command-line tools are also available to list at run-time the worlds
and their content ({\tt uwds ls}) or to display detailed informations on a
specific node ({\tt uwds show}).

@@

The {\tt uwds tf} client continuously monitor the ROS TF tree, and mirrors
TF frames as nodes in the desired \uwds world. A node is first created if none
matches a given TF frame, and its transformation is subsequently updated,
mirroring the TF frame. A regular expression can be provided to only mirror a
subset of the TF tree into \uwds.

Currently, the process is unidirectional: the {\tt uwds tf} client performs TF
to \uwds updates, but not the opposite.

@@

Spatial reasoning~\cite{O'Keefe1999} is a field in its own right, and has been
used for natural language processing for applications such as direction
recognition~\cite{Kollar2010,Matuszek2010} or language
grounding~\cite{Tellex2010}. Other examples in human-robot interaction include Ros
\etal~\cite{ros2010solving, ros2010which} which has been recently integrated
into a full architecture for autonomous human-robot
interaction~\cite{lemaignan2016artificial}.

@@

\uwds also implements an efficient algorithm to assess object visibility from a
specific viewpoint (\ie from a given \emph{perspective} node). The algorithm (color
picking) enables fast (single pass) computation of the visibility of every
object in the scene, while providing control regarding how many pixels should be
actually visible for the object to be considered globally visible. The
command-line tool {\tt uwds visibility} returns the list of visible objects from
the point of view of each camera in a given world, and \uwds also provides the
helper class {\tt VisibilityMonitor} to programmatically access visibility
information.

When integrated into a filter node, visibility computation allows to easily
create new worlds representing the estimated perspectives of the different
agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
\uwds client.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{l2tor-photo2}
    \caption{A user manipulating objects on top of a RFID mat (providing object
    localisation) with a NAO robot, while an operator monitors
    the interaction on a tablet that display an {\tt uwds view} client.}

@@

\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
in order to conceptualise and visualise the spatial relations and visibility of
the physical objects that participants interact with. In one of the scenarios of this
project, participants are instructed by a NAO robot to manipulate a number of
Duplo-like animals on top of a printed background paper which serves as a
landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
pedagogical strategies to promote spatial relation learning by object
manipulation. Such instructions include commands like to place the elephant
\textit{on top} or \textit{next to} the cube or put the giraffe \textit{inside}
the house.  To assess the performance, the robot and the operator need to
compute spatial relation between objects from the visual perspective of the
participant.

@@

The interaction scenario requires a number of \uwds clients. The client are
distributed across two computers (the robot and a Microsoft Windows-operated
tablet) that communicate with each other in order to synchronise the worlds
representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
the \uwds network, that consists in the following clients:

@@

\subsection{Relation to existing robotic middlewares}

Like traditional robotic middlewares, \uwds offers a form of distributed
computation based on message passing. However, it distinguishes itself from
existing middleware in several significant ways.

First and foremost, \uwds purposefully does not offer any \emph{general}
capability at distributing computation and data flows amongst independent
components: it focuses specifically on distributing environment models, both
spatial (geometric models) and temporal (events and situations).

While using standard middlewares as \emph{underlying transport} for \uwds would be
technically feasible and relatively easy to implement, it does not offer any
clear advantage over lighter and dedicated message passing libraries like ZeroMQ
or gRPC (the later being the one used by \uwds).

@@

We have presented as well a concrete instantiation of a system relying on \uwds
for its representation needs, and we have sketched future directions of
development.

>>>>

Software components accessing \uwds worlds are called clients. Some standard
clients (like a 3D visualisation tool) are provided with the \uwds package (see
section~\ref{std_clients}). Clients are otherwise written by the end users
using the \uwds client API (see section~\ref{api}).

@@

Four specific types of clients can be distinguished: \textbf{root clients} which
create and update worlds (`write-only' client, like the \emph{Sensor fusion} and
\emph{ROS interface} clients in Figure~\ref{fig|scene}); \textbf{leaf clients}
which on the contrary only read worlds, without modifying them (like the
\emph{Computation of spatial relations} client on Figure~\ref{fig|scene});
\textbf{filters} that copy an input world into an output world, performing
some filtering operation in-between (like the client \textit{Physics-based
position correction}); and \textbf{transformers} which transform one
representation into another (like the \emph{Semantic mapping} client on
Figure~\ref{fig|scene}).

@@

Each world is technically independent from all others. Dependencies between
worlds arise from the clients' connections. For instance, filters effectively
create a dependency between worlds. On Figure~\ref{fig|scene}, the \textit{Physics-based
position correction} client creates a dependency between the world {\tt base} (which
represents here the result of raw sensor fusion) and the world {\tt corrected}
which would be a physically-consistent copy of {\tt base}.
As a result, an \uwds network can be seen as a dependency graph between worlds (where
cyclic dependencies are permissible).

@@

Nodes in an \uwds scene graph have four possible types: \textbf{objects} that
represent concrete physical objects (typically with one or several associated 3D
meshes); \textbf{entities} that represent abstract entities like reference
frames or groups of objects; \textbf{perspectives} that represent viewpoints
of the scene (like cameras or human gaze); and \textbf{fields} that represent
scalar or vector fields (like the visibility of an object, the working space of
robot, \etc -- note that fields are not yet implemented in the current version
of \uwds, see section~\ref{futurework}).

@@

Complementing the spatial representation encapsulated in the scene graph, each
world also stores the world's \emph{timeline}. This data structure is shared
and synchronised amongst the clients in the same way as the scene graph.
Clients can record both \emph{events} (duration-less states) and
\emph{situations} in the timeline, \ie states with a start time and a (possibly open-ended) end time.

Importantly, the \uwds server automatically generates a snapshot of
the scene graph whenever an event or situation is added to the timeline. The
snapshot is associated to the event, which allows clients to effectively retrieve past
states of the world. This capability is anticipated to be primarily used by \uwds clients
performing action recognition.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{uwds-screenshot}
    \caption{Screenshot of the {\tt uwds view} 3D visualisation and manipulation
    tool. In this particular example, the 3D meshes have been pre-loaded using
    {\tt uwds load}. Their positions are then updated at run-time using the
    robot's sensors and proprioception (joint state).}
    \label{fig|uwds-view}
\end{figure}

@@

Critically, the network is essentially hidden to the client: from the user
perspective, the environment model is manipulated as a local data structure (see
Listing~\ref{lst|pythonapi}). Modifications to the model are asynchronously synchronised with
a central server (the {\tt underworlded} daemon) and broadcasted to every other
client connected to the same world.
% JK: underworlded - intentional? (I guess yes, but checking)

As previously mentioned, worlds are composite data structures comprised of a
scene graph and a timeline. These data structures are synchronised using
Google's gRPC message passing framework\footnote{\url{http://www.grpc.io/}}, ensuring
high throughput, reliability and cross-platform/cross-language support. The \uwds
API is specifically discussed hereafter, in section~\ref{api}.

\uwds is meant to broadcast complex environment representations (typically
including large geometric datasets, like meshes) in real-time. \uwds itself does
not perform many CPU intensive tasks (CPU intensive processing tasks -- sensor
fusion, physics simulation, \etc -- are performed by the clients themselves) and
as such, the performance bottleneck is essentially the network's data
throughput.  In that regard, one of the simple yet critical optimisations
performed by \uwds is automatic caching of mesh data. Mesh data are not
transmitted when nodes are updated; only a hash value of the mesh data. The
client can then request the full data whenever it is actually needed.

@@

As mentioned, \uwds uses Google's gRPC as message passing protocol. The protocol
is explicitly defined (using the \emph{protocol
buffers}\footnote{\url{https://developers.google.com/protocol-buffers/}}
interface definition language), and bindings to various languages and platforms
can be automatically generated from the protocol definition file (as of Feb
2017, gRPC can generate bindings for C, C++, C\#, Node.js, PHP, Ruby, Python, Go
and Java, on Windows, Mac and Linux).  The cross-platform/cross-language support
of gRPC is especially welcome in the academic context, as it offers ease and
flexibility to plug a variety of pre-existing components into an \uwds network.

@@

\uwds currently offers such a high-level client library for Python only (a C++
library is under development).  Listing~\ref{lst|pythonapi} gives a complete
example of an \uwds client performing simple filtering: the client continuously
listens for changes in an input world, removes some objects (in this case, items
whose name contain the string ``{\tt wall}''), and forwards all other changes to
an output world, effectively making the output world a copy of the input world
with all walls removed.

@@

        for node in in_world.scene.nodes:
            if("wall" not in node.name):
                out_world.scene.nodes.update(node)

\end{pythoncode}
    \caption{Example of a simple yet complete \uwds filter, written in Python:
    the client connects to the \uwds network, walks through the scene graph of
    world {\tt example\_in}, filters out some objects, and publishes the
    remaining objects in the world {\tt example\_out}.}

@@

Interestingly, while \uwds deals with 3D geometries and scenes, it does
represent 3D entities purely as data structures; no visual representation is
involved (and as such, the \uwds server and core libraries do not depend on any
graphics library like OpenGL). However, for all practical purposes, the ability
to visualise the content of a scene is desirable. \uwds provides a standard
client, {\tt uwds view}, that performs real-time 3D rendering of worlds,
using OpenGL (Figure~\ref{fig|uwds-view}).

This tool also supports basic object manipulations, that are broadcasted to the
other \uwds clients connected to the same world.

@@

Geometric data like meshes can be loaded into a world (or rather, into the whole
network as geometric data are treated as immutable resources shared amongst all
worlds) by simply sending an array of vertices and faces to the \uwds server.
As such, any client can load 3D data that is automatically made available to the
whole network. For example, an object recognition module recognises a new
object. It sends the mesh to the \uwds server and creates a new node which
references the mesh. At the next frame update, the 3D viewer sees a new node
referencing an unknown mesh. The viewer requests the mesh data from the
server, stores it internally (\eg as an OpenGL buffer), and renders it. Whenever
the node transformation is updated, the 3D viewer can re-render the node without
having to download the mesh from the server again.

Often, objects manipulated by the robot have known meshes with corresponding CAD
models that can be conveniently pre-loaded. In these cases, \uwds provides a
tool, {\tt uwds load}, that loads a mesh into a \uwds network (and optionally,
creates a node) from a large range of 3D formats (including Collada, FBX, OBJ,
Blender)\footnote{The underlying import
capability is provided by the {\sc assimp} library.
\url{http://assimp.sourceforge.net/}}.

@@

\uwds provides a range of tools to inspect a running network. Graphical
tools ({\tt uwds explorer} and {\tt uwds timeline}, see Figure~\ref{fig|explorer})
provide a user-friendly overview of the system's graph with the connections
between the clients and the worlds, as well as their activity.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{tools}
    \caption{Screenshots of the timeline (top) and network (bottom)
    introspection tools. Worlds are represented as boxes; clients as ellipses.}
    \label{fig|explorer}
\end{figure}

Specialised command-line tools are also available to list the worlds
and their content ({\tt uwds ls}) at run-time, or to display detailed information for a
specific node ({\tt uwds show}).

@@

The {\tt uwds tf} client continuously monitors the ROS TF tree, and mirrors
TF frames as nodes in the desired \uwds world. A node is first created if none
matches a given TF frame, and its transformation is subsequently updated,
mirroring the TF frame. A regular expression can be provided to only mirror a
subset of the TF tree into \uwds.

Currently, the process is unidirectional: the {\tt uwds tf} client performs TF
to \uwds updates, but not the reverse.

@@

Spatial reasoning~\cite{O'Keefe1999} is a field in its own right, and has been
used for natural language processing for applications such as direction
recognition~\cite{Kollar2010,Matuszek2010} or language
grounding~\cite{Tellex2010}. Other examples in human-robot interaction include Ros
\etal~\cite{ros2010solving, ros2010which} which has recently been integrated
into a full architecture for autonomous human-robot
interaction~\cite{lemaignan2016artificial}.

@@

\uwds also implements an efficient algorithm to assess object visibility from a
specific viewpoint (\ie from a given \emph{perspective} node). The algorithm (color
picking) enables fast (single pass) computation of the visibility of every
object in the scene, while providing control regarding how many pixels should actually be
visible for the object to be considered globally visible. The
command-line tool {\tt uwds visibility} returns the list of visible objects from
the point of view of each camera in a given world, and \uwds also provides the
helper class {\tt VisibilityMonitor} to programmatically access visibility
information.

When integrated into a filter node, visibility computation allows easy
creation of new worlds representing the estimated perspectives of the different
agents. Listing~\ref{lst|humanperspective} provides the complete code of such an
\uwds client.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{l2tor-photo2}
    \caption{A user manipulating objects on top of an RFID mat (providing object
    localisation) with a NAO robot, while an operator monitors
    the interaction on a tablet that displays an {\tt uwds view} client.}

@@

\uwds has been used within the European L2TOR project~\cite{belpaeme2015l2tor}
in order to conceptualise and visualise the spatial relations and visibility of
the physical objects that participants interact with. In one of the scenarios of this
project, participants are instructed by a NAO robot to manipulate a number of
Duplo-like animals on top of a printed background paper which serves as a
landscape (Figure~\ref{fig|l2torexample}). The robot instructions are based on
pedagogical strategies to promote spatial relation learning by object
manipulation. Such instructions include commands to place the elephant
\textit{on top} or \textit{next to} the cube, or put the giraffe \textit{inside}
the house.  To assess the performance, the robot and the operator need to
compute spatial relationships between objects from the visual perspective of the
participant.

@@

The interaction scenario requires a number of \uwds clients. The clients are
distributed across two computers (the robot and a Microsoft Windows-operated
tablet) that communicate with each other in order to synchronise the world
representations. Figure~\ref{fig|l2torarchitecture} illustrates the topology of
the \uwds network, that consists in the following clients:

@@

\subsection{Relation to existing robotic middleware}

Like traditional robotic middleware, \uwds offers a form of distributed
computation based on message passing. However, it distinguishes itself from
existing middleware in several significant ways.

First and foremost, \uwds purposefully does not offer any \emph{general}
capability to distribute computation and data flow amongst independent
components: it focusses specifically on distributing environment models, both
spatial (geometric models) and temporal (events and situations).

While using standard middleware as \emph{underlying transport} for \uwds would be
technically feasible and relatively easy to implement, it does not offer any
clear advantage over lighter and dedicated message passing libraries like ZeroMQ
or gRPC (the later being the one used by \uwds).

@@

We have additionally presented a concrete instantiation of a system relying on \uwds
for its representation needs, and we have sketched future directions of
development.
